---
title: "Untitled"
author: "Mireille Vargas"
date: "5/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(rgee)
library(mapview)
library(exactextractr)
library(raster)
library(sf)
library(tidyverse)
library(dplyr)
library(tigris)
library(leaflet)
library(sp)
library(rgeos)
library(geojsonio)
library(googledrive)
library(stars)
```


```{r, echo=FALSE}
ee_Initialize()
ee_check()

# ee_check_python()
# ee_check_credentials()
# ee_check_python_packages()
```

Grab the area of interest
```{r}
counties <- ee$FeatureCollection("TIGER/2016/Counties")
myfilter <- ee$Filter$inList(
  opt_leftField = "NAME",
  opt_rightValue = list(
    "Polk"
  )
)

polk <- counties$filter(myfilter)
```


```{r}
years_all_gridmet <- data.frame(
  start = c((seq(as.Date("1979-01-01"), as.Date("2021-03-01"), by = "month"))) %>% 
    as.character() %>% 
   append( c((seq(as.Date("1979-01-15"), as.Date("2021-03-15"), by = "month"))) %>% 
    as.character()),
  end =   c((seq(as.Date("1979-01-14"), as.Date("2021-03-14"), by = "month"))) %>% 
    as.character() %>%
    append(c((seq(as.Date("1979-02-01"), as.Date("2021-04-01"), by = "month")) - 1) %>% 
    as.character()),
  counts = c(seq("1", "1014"))
)
```

```{r}
bartow <- ee$Geometry$Point(c(27.88608, -81.83251)) #latitude goes first and then longitude

aoi <- ee$Geometry$Polygon(
  coords = list(
    c(27.84601, -81.91396),
    c(27.84601, -81.77114),
    c(27.94736, -81.77114),
    c(27.94736, -81.91396),
    c(27.84601, -81.91396)
  )
)

#create an sf object
DT <- data.frame(
  lat1 = 27.88608,
  long1 = -81.83251
)

aoi2 <- st_as_sf(DT, coords = c("long1", "lat1"), remove = FALSE) %>% 
  st_set_crs(4326)
```

```{r}
#check

landsat <- ee$Image("LANDSAT/LC08/C01/T1_TOA/LC08_044034_20140318")
vizParams <- list(
  bands = c('B5', 'B4', 'B3'),
  min = 0,
  max = 0.5,
  gamma = c(0.95, 1.1, 1)
)
Map$setCenter(lon =-81.83251, lat = 27.88608, 10)
Map$addLayer(landsat, vizParams, 'false color composite')
```

Create functions that grab each band rename it

Grab bands. The bands of interest are:
pr: precipitation
rmax: max relative humidity
rmin : min relative humidity
tmmn : min temperature
tmmx: max temperature
srad : Surface downward shortwave radiation
vs: Wind velocity at 10m

```{r}
bands <- c("pr", "rmax", "rmin", "tmmn", "tmmx", "srad", "vs")
```

test
```{r}
  gridmet<- ee$ImageCollection("IDAHO_EPSCOR/GRIDMET")$ #grabbing maximum relative temperature from GridMET
  filterBounds(aoi)$ #filtering within the bay area counties
  filterDate(years_all_gridmet$start[1], years_all_gridmet$end[1])$
  select(bands)$ #all interested features
  mean()
# trying out extract
   values <- ee_extract(
   x = gridmet,
   y = aoi2["lat1"],
   scale = 4000,
   fun = ee$Reducer$mean(),
   sf = TRUE
  ) #Error in data.frame(..., check.names = FALSE) : arguments imply differing number of rows: 1, 0

#trying out the method from stack exchange
yieldLocation <- ee$FeatureCollection(table) #giving me issues
result <- gridmet$reduceRegions(yieldLocation, ee$Reducer$mean(), 1)
```



Going through the whole loop
```{r}
for (i in 1:nrow(years_all_gridmet)){
  gridmet<- ee$ImageCollection("IDAHO_EPSCOR/GRIDMET")$ #grabbing maximum relative temperature from GridMET
  filterBounds(bartow)$ #filtering within the bay area counties
  filterDate(years_all_gridmet$start[i], years_all_gridmet$end[i])$
  select(bands)$ #all interested features
  median()
  
   values <- ee_extract(
   x = gridmet,
   y = aoi2["lat1"],
   scale = 4000,
   fun = ee$Reducer$mean(),
   sf = TRUE
  )
  
  #create unique dataframes to later bind
  assign(paste0("values_", years_all_gridmet$counts[i], sep = ""), values)
  
  #Save! Especially if you have bad wifi!!!!!!
  saveRDS(values, paste0("values_", years_all_gridmet$counts[i], ".rds"))
  
  #print for your own sanity
  print(i)
  
}
```

Would then bind all the values together