---
title: "Rain Gauge"
author: "Mireille Vargas"
date: "5/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(lubridate)
```


```{r}
# #read in data
# 
# rain <- read.csv("D:/NOAA_rain_gauge/Rain_Gauge_summary.csv")
# 
# rain_vals <- rain$Prec_no_na
```



```{r}
# years_all <- data.frame(
#   start = c((seq(as.Date("1979-01-01"), as.Date("2021-03-01"), by = "month"))) %>% 
#     as.character() %>% 
#    append( c((seq(as.Date("1979-01-15"), as.Date("2021-03-15"), by = "month"))) %>% 
#     as.character()),
#   end =   c((seq(as.Date("1979-01-14"), as.Date("2021-03-14"), by = "month"))) %>% 
#     as.character() %>%
#     append(c((seq(as.Date("1979-02-01"), as.Date("2021-04-01"), by = "month")) - 1) %>% 
#     as.character())
# )
# 
# 
# data <- years_all[order(years_all$start),]
# 
# data <- data %>% 
#   cbind(rain_vals)
# 
# #write.csv(data, "rain_gauge_data.csv")
```

Read in the data
```{r}
# rain_gauge_data <- read.csv("rain_gauge_data.csv") %>% #precipitation is in inches
#   select(
#     start,
#     end,
#     rain_vals
#   )
# 
# #organize the gridmet data from bartow
# gridmet_bartow <- readRDS("gridmet_vals_bartow.rds") %>% 
#   dplyr::select(!counts)
# 
# 
# gridmet_bartow <- gridmet_bartow[order(gridmet_bartow$start),]
# 
# sum(rain_gauge_data$rain_vals == 0.00, na.rm = TRUE)
# sum(is.na(rain_gauge_data))
# 
# #look at na values
# new_DF <- rain_gauge_data[is.na(rain_gauge_data$rain_vals),]
# 
# no_na <- rain_gauge_data[!is.na(rain_gauge_data$rain_vals),]
# zero_df <- no_na[(no_na$rain_vals == 0.00),]
# 
# #filter
# #bartow <- polk_county[polk_county$NAME %in% "BARTOW",]
# 
# gridmet_bartow_na <- gridmet_bartow[gridmet_bartow$start %in% new_DF$start,] #there are 3 zero values!
# 
# gridmet_bartow_zero <- gridmet_bartow[gridmet_bartow$start %in% zero_df$start,]
```

This is the actual cleaning. Focus here.

This chunk selects the start, end, and rain_vals (sum of rain in inches for time range). I then seperate the dates so I can calculate how many days lapsed in each time range. Then I create a new column of average precipitation in mm based on the difference in the days in the time range and converting inches to mm. I join in the gridmet data and select only my desired columns. Then i transpose my dataframe and convert back to a dataframe.
```{r}
#clean up data to remove NAs and to convert inches to mm
rain_gauge_data <- read.csv("rain_gauge_data.csv") %>% #precipitation is in inches
  select(
    start,
    end,
    rain_vals
  )

gridmet_bartow <- readRDS("gridmet_vals_bartow.rds") %>% 
  dplyr::select(!counts)

all_data_bartow <- read.csv("rain_gauge_data.csv") %>% #precipitation is in inches
  select(
    start,
    end,
    rain_vals
  )

all_data_bartow <- all_data_bartow %>%
  separate(start, into = c('year_start', 'month_start', 'day_start')) %>%
  
  separate(end, into = c('year_end', 'month_end', 'day_end'))

all_data_bartow$day_start <- all_data_bartow$day_start %>% 
  as.integer()
all_data_bartow$day_end <- all_data_bartow$day_end %>% 
  as.integer()  
  
all_data_bartow <- all_data_bartow %>% 
  mutate(
    rain_avg_mm = (rain_vals * 25.4)/(day_end - day_start)
  ) %>% 
  mutate(
    rain_avg_in = rain_vals/(day_end - day_start)
  ) %>% 
  mutate(
    start = rain_gauge_data$start
  ) %>% 
  left_join(
    gridmet_bartow,
    by = "start"
  ) %>% 
  select(
    #rain_avg_mm,
    #rain_avg_in,
    start,
    end, 
    rmax,
    rmin,
    srad,
    tmmn,
    tmmx,
    vs
  ) %>% 
  t() %>% 
  as.data.frame()

```

Creating the Y.dat of previous time steps of precipitation. First I had to recreate some of the steps from the previous chunk to grab average precipitation in mm. Then I added that column into my y dat dataframe. From there I created 10 new columns grabbing the previous 10 time steps. I deselected the first ten rows that would ahve NA values. I then transposed the rows and columns and converted to a dataframe.
```{r}
#library(data.table)
rain_avg <- read.csv("rain_gauge_data.csv") %>% #precipitation is in inches
  select(
    start,
    end,
    rain_vals
  )

rain_avg <- rain_avg %>%
  separate(start, into = c('year_start', 'month_start', 'day_start')) %>%
  
  separate(end, into = c('year_end', 'month_end', 'day_end'))

rain_avg$day_start <- rain_avg$day_start %>% 
  as.integer()
rain_avg$day_end <- rain_avg$day_end %>% 
  as.integer()  
  
rain_avg <- rain_avg %>% 
  mutate(
    rain_avg_mm = (rain_vals * 25.4)/(day_end - day_start)
  ) %>% 
  mutate(
    rain_avg_in = rain_vals/(day_end - day_start)
  )

rain_avg_mm <- rain_avg$rain_avg_mm

y_dat <- rain_gauge_data %>% 
  cbind(rain_avg_mm) %>% 
  select(start,
         end,
         rain_avg_mm)

y_dat <- y_dat %>% 
  mutate(
    prev_1 = lag(
      rain_avg_mm, 1
    )
  ) %>% 
  mutate(
    prev_2 = lag(
      rain_avg_mm, 2
    )
  ) %>% 
  mutate(
    prev_3 = lag(
      rain_avg_mm, 3
    )
  ) %>% 
  mutate(
    prev_4 = lag(
      rain_avg_mm, 4
    )
  ) %>% 
  mutate(
    prev_5 = lag(
      rain_avg_mm, 5
    )
  ) %>% 
  mutate(
    prev_6 = lag(
      rain_avg_mm, 6
    )
  ) %>% 
  mutate(
    prev_7 = lag(
      rain_avg_mm, 7
    )
  ) %>% 
  mutate(
    prev_8 = lag(
      rain_avg_mm, 8
    )
  ) %>% 
  mutate(
    prev_9 = lag(
      rain_avg_mm, 9
    )
  ) %>% 
  mutate(
    prev_10 = lag(
      rain_avg_mm, 10
    )
  )

#deselect the first ten rows
y_dat <- y_dat[c(11:1014),]
  
#now transpose and create into a data frame
y_dat <- y_dat %>% 
  t() %>% 
  as.data.frame()
```

Save the data as csv files
```{r}
write.csv(all_data_bartow, "x_data.csv")
write.csv(y_dat, "y_data.csv")
```

